Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2019-02-28 review-responses.tex

    Overall, I think this article makes a very valid contribution for Molecular
    Ecology Ressources. The authors put their work into the right context and
    show good knowledge of the relevant literature. The writing is both clear
    and concise. Importantly, they do not shy away from pointing out the limits
    of their method, and the very stringent (often unrealistic) assumptions
    such methods usually make.

    I have several comments, but none of them are ``make-or-break'' or fatal flaws.
    The manuscript is already in a good state. But I hope the authors can use my
    comments to further improve their work.  
\end{quote}

\begin{point}{}
    Clearer point out the current computational limitations of your coalescence
time method. It seems it gets tricky with more than about 30 demes? What is the
run-time then, does the whole thing blow up? Give an idea how big is too big!
\end{point}

\reply{
    The runs with 16 demes take around 10 hours with the number of iterations we used.
    30 is probably a good limit for the current implementation. 
    There are changes that could be made to make it more efficient for larger systems, 
    but at that point the ill conditioning problem would be more severe 
    because there would be more parameters, 
    so we would probably have to do something similar to EEMS, 
    where movement rates are the same within regions, 
    but in a way that still allowed asymmetric migration.
}

@@

\begin{point}{}
    What is the exact scaling? The bottleneck seems to be calculating the
    coalescence times with equation (7). You only say that it ``can be solved
    with standard linear algebra''. Are you doing Gaussian elimination or
    something else worse than $N^2$? Do you use any speed-ups to
    deal with the sparse nature of G for neighbor-only migration? Is it faster
    than getting the coalescence times by iterating time-discretized
    recursions?  It would be helpful for further progress to describe the exact
    scaling and the exact algorithm you use.
\end{point}

\reply{
    In the current implementation, 
    equation (7) is solved using the solve(A,b) in R. 
    The actual method varies depending on the specific linear algebra library used, 
    but it is typically Gaussian elimination or LU decomposition, 
    both of which are $\mathcal{O}(N^3)$ 
    (unfortunately the decomposition is different each step;
    if it were the same it would be $\mathcal{O}(N^2)$).
    Since $n$ demes gives you order $n^2$ equations and unknowns, 
    the time complexity is 
}

@@

\begin{point}{}
    You use mean absolute error in movement rate to describe the model fit. I
    believe this measure can be quite misleading for exponential movement rates.
    Large rates can be notoriously hard to infer (are you super or super-super well
    connected is almost the same), so it is easy to be off quite a bit, which
    drives the average absolute error in rate. Also in practice, one does not care
    so much about absolute rates but ``relative'' ones. So using a correlation
    measure R**2  of inferred movement rates with actual movement rates would be a
    very useful and natural fit parameter.  
\end{point}

\reply{
}

\begin{point}{}
    You write you use standard methods to check convergence and mixing of the
    MCMC chain (line 182). Given the degeneracy and the high dimensionality,
    where there any problems? E.g., local optima you were stuck sometimes? Or
    was it always pretty straightforward? Maybe write a few sentences on that!
\end{point}

\reply{
    Convergence could be slow, and the likelihood surface is often ridgy, 
    but we did not see too many problems with getting stuck in completely wrong
    areas of state space.
    This is partly how ``pre-burn-in'' helps.
}

@@

\begin{point}{}
    $G$, the movement rate, is clearly a key parameter but is not clearly defined (or
    if it was I missed it).  What are the units of G?  Geographic distance per
    unit time?  What are the time units?  Years?  Generations?  If the latter,
    then does G represent parent-offspring dispersal distance as defined by
    Wright?  Are generations assumed to be discrete?  If not, can a migrant
    leave offspring at various different geographic points?
\end{point}

\reply{
}

\begin{point}{}
    As described, the model deals with groups of individuals (populations) that
    sometimes exchange individuals.  But much of landscape genetics/genomics
    deals with continuously-distributed species that don’t have recognizable
    populations.  Are the results also applicable to scenarios like this?  Why
    or why not?
\end{point}

\reply{
    Yes, this is one of the reasons that continuous space was used 
    in the forwards time simulations. 
    As long as the landscape behaves as if it is fairly continuous
    and changing the boundary does not move a large number of individuals 
    from one side to the other, 
    the exact boundary in the discretization should not have a large effect.
    Putting two isolated subpopulations into a single group would also be problematic.
}

@@

\begin{point}{}
    ``Can our method replace resistance-based methods? Perhaps, but the substantial
    uncertainty we saw on graphs with only tens of nodes is indicative of a
    larger problem we face for realistic models. We have seen that
    discretization of space results in substantial modeling error, in part
    because of randomness of geographic sampling and unmodeled process noise
    can lead to overfitting. Partitioning space into a finer grid should help
    with these problems, but tends to make the inference problem itself more
    ill-conditioned: with more connections, changing the value of one
    connection affects coalescence time less, and so inferences about that
    value must necessarily be less certain. This tradeoff implies some degree
    of unavoidable uncertainty.  Reproducible, reliable inference will likely
    require development of new inference methods that explicitly model
    continuous geography. `` This seems like a fairly crucial paragraph.  But
    what exactly is the take-home message for biologists interested in
    practical application of the methodology?
\end{point}

\reply{
}

\begin{point}{}
    Coalescent rates are critical to model performance, which can be poor if the
    rate is too low.  It is obvious that the coalescent rate is inversely
    related to Ne, but in practical terms, what does it mean when the
    coalescent rate is lowered from 10 to 0.1?
\end{point}

\reply{
    Assuming that the movement parameters are of order one, 
    a coalescence rate of 10 means that 
    lineages would be more likely to coalesce the first time they are in the same location
    and a coalescence rate of 0.1 means that lineages would likely meet many times
    before finally coalescing. 
    For example, if the gene flow rates out of a deme for both lineages sum to 8
    (rate of 1 for each direction for each lineage), 
    the chance of the lineages coalescing before one of them leaves 
    would be 10/18 (~56\%) with a coalescence rate of 10 
    but only 0.1/8.1 (~1.2\%) with a coalescence rate of 0.1.
}

@@

\begin{point}{\revref}
    ``if fecundity is variable'' -- on what scale?  Among populations?  Among
    individuals?  Among ages within the same individual? 
\end{point}

\reply{
}

\begin{point}{\revref}
    ``It is not required in this approach to have samples from every spatial
    location ...''  Required for what?  Are they saying that unsampled but real
    (ghost) populations have no effect on precision or bias?
\end{point}

\reply{
    In practice, the more locations with samples you have, the better the result will be.
    We would not typically recommend doing inference that includes unsampled demes
    if it could be avoided.
    However, in the theoretical case of a population fitting the model perfectly
    and knowing the other genetic distance values perfectly, 
    it is possible to solve exactly for paramters relating to the unsampled populations
    as long as there are as many unique equations as unknowns.
    In a real situation, if data is very good for the sampled locations, 
    inference should be reasonable for the unsampled ones.
}

### 2019-02-28 isolation_by_coalescence.tex
more review responses

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed. \revpoint{3}{2}
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
This implies a fundamental limit to the geographical resolution of inferred maps.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.

>>>>

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed \revpoint{3}{2}
(i.e., an individual is unlikely to live many demes away from their parents).
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
This implies a fundamental limit to the geographical resolution of inferred maps.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.
